Theory about Normal Basis on $F(2^k)$

### 2017-03-03 dissertation/submission/abstract.tex
dissertation after defense update 2
after submission: revision 2

\begin{center}{\bf ABSTRACT}\end{center}
Formal verification of hardware designs has become an essential
component of the overall system design flow. The designs are generally
modeled as finite state machines, on which property and equivalence
checking problems are solved for verification. Reachability analysis
forms the core of these techniques. However, increasing
size and complexity of the circuits causes the state explosion
problem. Abstraction is key to tackle the scalability challenge.

@@

While algebraic geometry provides powerful abstraction and reasoning
capabilities, the algorithms exhibit high computational complexity. In
the dissertation, we show that by analyzing the constraints, it is
possible to obtain more insights about the polynomial ideals, which can be
exploited to overcome the complexity. Using our algorithm design and
implementations, we demonstrate how to perform reachability
analysis of finite-state machines purely at the word-level. Using this
concept, we perform scalable verification of sequential arithmetic
circuits. As contemporary approaches make use of resolution proofs and
unsatisfiable cores for state-space abstraction, we introduce the
algebraic geometry analog of unsatisfiable cores, and present
algorithms to extract and refine unsatisfiable cores of polynomial
ideals. Experiments are performed to demonstrate the efficacy of our
approaches.

>>>>

%\begin{center}{\bf ABSTRACT}\end{center}
Formal verification of hardware designs has become an essential
component of the overall system design flow. The designs are generally
modeled as finite state machines, on which property and equivalence
checking problems are solved for verification. Reachability analysis
forms the core of these techniques. However, increasing
size and complexity of the circuits causes the state explosion
problem. Abstraction is the key to tackling the scalability challenges.

@@

While algebraic geometry provides powerful abstraction and reasoning
capabilities, the algorithms exhibit high computational complexity. In
the dissertation, we show that by analyzing the constraints, it is
possible to obtain more insights about the polynomial ideals, which can be
exploited to overcome the complexity. Using our algorithm design and
implementations, we demonstrate how to perform reachability
analysis of finite-state machines purely at the word level. Using this
concept, we perform scalable verification of sequential arithmetic
circuits. As contemporary approaches make use of resolution proofs and
unsatisfiable cores for state-space abstraction, we introduce the
algebraic geometry analog of unsatisfiable cores, and present
algorithms to extract and refine unsatisfiable cores of polynomial
ideals. Experiments are performed to demonstrate the efficacy of our
approaches.

### 2017-03-03 dissertation/submission/acknowledge.tex
after submission: revision 2

I was born in a small town in mid-inland-China. My grandpa, Shouli Sun, 
is the first man to educate me. He was so kind and patient, and his education 
was so successful that I could read 2,000 Chinese characters at the age of 3, 
when my grandpa passed away. With a good start, I became used to being the 
first place in a variety of exams throughout elementary school, middle school, 
and high school. Peking university seemed to be the end of my personal 
progress in education, until I learnt that there is a way to continue my 
graduate study abroad, more specifically, in the US. So I would like to 
thank Dr. Fu, Mr. Qiu and Mr. Yang for providing such information and 
suggestions as my roommates at that time.

The beginning of my graduate school experience was not 
so wonderful. I got lost among the labs of the ECE department since there are 
too many directions and categories of research. Finally, I found out what
I am really interested in after taking the lectures by my advisor, 
Prof. Priyank Kalla. He led the way towards designing a beautiful and powerful
algorithm and applying it to the ``mission impossible" -- verifying
a large circuit with 100\% fault coverage. I would like to thank him for
his valuable guidance, magnificent inspirations, and of course, his generous 
financial support that allow me to carry out my research. Meanwhile, I would 
like to thank four professors serving as my supervisory committee members, 
for their kind help in assisting my research and coursework, as well as the 
review of this dissertation.

>>>>

I was born in a small town in mid-inland-China. My grandpa, Shouli Sun, 
was the first man to educate me. He was so kind and patient, and his education 
was so successful that I could read 2,000 Chinese characters at the age of 3, 
when my grandpa passed away. With a good start, I became used to taking the 
first place in a variety of exams throughout elementary school, middle school, 
and high school. Peking university seemed to be the end of my personal 
progress in education, until I learnt that there is was way to continue my 
graduate study abroad, more specifically, in the US. So I would like to 
thank Dr. Fu, Mr. Qiu and Mr. Yang for providing such information and 
suggestions as my roommates at that time.

The beginning of my graduate school experience was not 
so wonderful. I got lost among the labs of the ECE department since there are 
too many directions and categories of research. Finally, I found out what
I am really interested in after taking the lectures by my advisor, 
Prof. Priyank Kalla. He led the way towards designing a beautiful and powerful
algorithm and applying it to the ``mission impossible" -- verifying
a large circuit with 100\% fault coverage. I would like to thank him for
his valuable guidance, magnificent inspirations, and of course, his generous 
financial support, which allow me to carry out my research. Meanwhile, I would 
like to thank four professors serving as my supervisory committee members, 
for their kind help in assisting my research and coursework, as well as the 
review of this dissertation.

### 2017-03-03 dissertation/submission/appendix1.tex
after submission: revision 2

\chapter{Normal Basis Theory}
\label{append:NB}
In Chapter \ref{ch:prelim_GF} we briefly introduced the concept of normal basis (NB), and 
the benefits of using NB. In this appendix chapter we describe more details about NB theory by 
characterizing NB from a linear algebra perspective, constructing a general NB in an 
arbitrary field, and converting between NB and StdB. All theorems and lemmas refer to the 
dissertation of Gao \cite{gao:phd_normal_basis}, and we deduce all the proofs for them.
\section{Characterization of Normal Basis}
In order to depict the characterization of NB, we need the introduce some concepts from linear algebra domain.
\begin{Definition}[Frobenius Map]
Define map $\sigma:x\to x^p,~x\in \F_{p^n}$. This map denoting the linear map of field extension $\F_{p^n}$
over $\F_p$.
\end{Definition}
Additionally we can define a subspace based on linear map $T$:
\begin{Definition}
A subspace $W\subset V$ is called {\bf $T$-invariant} when 
$$Tu \in W, ~~ \forall \text{ vector }u\in W$$

@@

We define the nullspace of a polynomial as:
\begin{Definition}[Nullspace of a polynomial]
\label{def:nullspace}
For any polynomial $g(x) \in \F[x]$, the null space of $g(T)$ consists of all vectors $u$ such that
$g(T)u = 0$.
\end{Definition}
Finally we get to the most important concept: $T$-Order. It derives the construction of a ring extended 
by a field requiring a minimal polynomial.
\begin{Definition}
For any vector $u \in V$, the monic polynomial $g(x) \in \F[x]$ with smallest degree such that
$g(T)u = 0$ is called the {\bf $T$-Order} of $u$ or {\bf minimal polynomial} of $u$.

@@

Using the concepts introduced above, along with other basic concepts in linear algebra and finite field theory,
we can derive following theorems.

@@

Let $k$ be a polynomial whose degree is larger than that of $f(x)$, {\it i.e.} $deg(k(x))>deg(f(x))$, then
$$f(x)|k(x)~~\text{iff}~k(T)=0$$

@@

This equation denotes that $\beta$ is one of $p^n$ roots of polynomial $\Func(x) = \sum_if_i(x^{p^i})$.
However, the maximum number of roots allowed equals to the degree of $\Func(x)$, which is 
$p^{n-1}<p^n$, we find a contradiction. Thus we conclude that
$$\textit{Both characteristic and minimal polynomial is }x^n-1$$
\end{Proof}

From lemmas above we can deduce the following corollary:
\begin{Corollary}
\label{corol:1}
An element $\alpha \in \F_{p^n}$ is a {\bf normal element} if and only if $Ord_{\alpha,\sigma}(x) = x^n-1$.
\end{Corollary}
\begin{Proof}
Normal bases require $\{\beta,\beta^2,\dots,\beta^{2^{n-1}}\}$ to be linearly independent,
which is equivalent to the following expression:
$$\forall f(x)\in \F_p[x],~deg(f(x))<n$$
Furthermore, because there is no annihilators in the $T$-Order, this implies that
$$Ord_{\alpha,\sigma}(x) = x^n-1$$
\end{Proof}

@@

Ultimately, since there always exist at least one element fulfilling requirements in Theorem \ref{thm:1}
or Theorem \ref{thm:2}, we obtain following theorem:
\begin{Theorem}[Normal Basis Theorem over Finite Fields]
There always exists a normal basis of $\F_{p^n}$ over $\F_p$.
\end{Theorem}
\section{Construction of General Normal Bases}
After proving the existence of normal bases, the upcoming question is that how to find such a normal basis/element.
For general normal basis identification, there are two methods widely used: the L\"uneburg's algorithm
and Lenstra's algorithm.

\subsection{L\"uneburg's Algorithm}
L\"uneburg's algorithm can be described with following steps:

@@

Use Lemma \ref{lem:3}, the minimal/characteristic polynomial for $\sigma$ is $x^n-1$ implies that
any annihilator of $\alpha^i$ ({\it i.e.} $f_i(\sigma)$) divides $x^n-1$. Meanwhile, $\{\alpha^i~|~0<i<n-1\}$ forms a (standard) basis
of field $\F_{p^n}$, this means $\alpha_i$ are linearly independent with each other ($i=0,1,\dots,n-1$).
Since linear map $f(\sigma)$ guarantees that all elements in the field 
have order $lcm(f_0,\dots,f_{n-1})$, {\it i.e.} $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
By contradiction we can prove that there is no factor of $x^n-1$ that can be divided by the product 
of elements in set $\{f_0,f_1,\dots,f_{n-1}\}$. This actually corresponds to the assertion in the first step of the algorithm:
$$x^n-1 = lcm(f_0,f_1,\dots,f_{n-1})$$

@@

\subsection{Lenstra's Algorithm}
L\"uneburg's algorithm is an analytic method, which requires relatively high computational complexity 
mainly because of the factorization in its second step. To overcome the high cost, an inductive method 
is proposed, which is the Lenstra's algorithm. It allows for setup of heuristics to accelerate the procedure.
Before introducing the details of the algorithm, we demonstrate two preliminary lemmas.
\begin{Lemma}
\label{lem:4}
For an arbitrary element $\theta \in \F_{p^n}$ that $Ord_\theta(x) \neq x^n - 1$,
let $g(x) = (x^n - 1)/Ord_\theta(x)$. There exists another element $\beta$ such that $g(\sigma)\beta = \theta$.
\end{Lemma}
\begin{Proof}
Assume $\gamma$ is the desired normal element. From the definition of normal element we derive 
$$\exists f(x)\in \F_{p^n}[x],~~f(\sigma)\gamma = \theta$$

@@

\begin{Lemma}
\label{lem:5}
Define $\theta$ and $g(x)$ as in Lemma \ref{lem:4}. Assume there exists a solution 
$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Respectively there exists a non-zero element $\eta$ such that
$$g(\sigma)\eta = 0$$
and
$$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$$
\end{Lemma}
\begin{Proof}
Assume $\gamma$ is the desired normal element. Similarly,
$$\exists \eta = Ord_\theta(\sigma)\gamma \neq 0,~~g(\sigma)\eta = 0$$

@@

Based on Lemmas \ref{lem:4} and \ref{lem:5}, the Lenstra's algorithm is described below:

\begin{enumerate}[{1)}]
\item Take an arbitrary element $\theta \in \F_{p^n}$, determine $Ord_\theta(x)$.
\item If $Ord_\theta(x) = x^n - 1$ then algorithm terminates and return $\theta$ as a normal element.
\item Otherwise, compute $g(x) = (x^n - 1)/Ord_\theta(x)$, and solve $\beta$ from $g(\sigma)\beta = \theta$.
\item Determine $Ord_\beta(x)$. If $deg(Ord_\beta(x)) > deg(Ord_\theta(x))$ then replace $\theta$ by $\beta$ and go to 2);
otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a non-zero element $\eta$ such that $g(\sigma)\eta = 0$,
replace $\theta$ by $\theta + \eta$ and determine the order of new $\theta$, and go to 2).
\end{enumerate}

@@

Directly constructing $\lambda$-Matrix is difficult for general NBs. 
In Section \ref{sec:nbdesign}, we introduced the concept of multiplication table (M-table) and explained how its 
entries are transformed from the $0$-th $\lambda$-Matrix. 
Recall the M-table is defined as
\begin{equation}
\label{eqn:appmt}
\beta
\begin{bmatrix}
\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
\end{bmatrix}
= {\bf T}
\begin{bmatrix}
\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
\end{bmatrix}
\end{equation}
Therefore we can compute M-table and then covert it to $\lambda$-Matrix.
Using Equation \ref{eqn:appmt}, we derive entry $T_{i,j}$ in the M-table by equating it to
the $j$-th bit of the NB representation of element $\beta\cdot\beta^{2^{i}}$.
For example, the $i$-th row of the M-table for NB in Example \ref{ex:Rijndael} satisfies
$$\beta\cdot\beta^{2^i} = \sum_{j=0}^{7} T_{i,j}\beta^{2^j}$$
Thus, the M-table can be written as
\begin{equation*}
T = \left(\begin{array}{lccccccr}
0 &1 &0 &0 &0 &0 &0 &1 \\
0 &1 &0 &0 &1 &1 &0 &1 \\
0 &0 &1 &0 &0 &1 &0 &1 \\
0 &0 &0 &0 &0 &0 &0 &1 \\
1 &1 &1 &1 &1 &1 &1 &1 \\
0 &0 &0 &0 &1 &0 &0 &0 \\
1 &1 &0 &1 &0 &1 &0 &0 \\
1 &0 &0 &1 &1 &0 &1 &0
\end{array}\right)
\end{equation*}
Use equation $M_{i,j}^{(0)} = T_{j-i,-i}$, we can obtain the
$\lambda$-Matrix in Equation \ref{eqn:mat_R}.

>>>>

\chapter{Normal Basis Theory}
\label{append:NB}
In Chapter \ref{ch:prelim_GF} we briefly introduced the concept of normal basis (NB), and 
the benefits of using NB. In this appendix chapter we describe more details about NB theory by 
characterizing NB from a linear algebra perspective, constructing a general NB in an 
arbitrary field, and converting between NB and StdB. All theorems and lemmas refer to the 
dissertation of Gao \cite{gao:phd_normal_basis}, and we deduce all the proofs for them.
\section{Characterization of Normal Basis}
In order to depict the characterization of NB, we need to introduce some concepts from the domain of linear algebra.
\begin{Definition}[Frobenius Map]
Define map $\sigma:x\to x^p,~x\in \F_{p^n}$. This map denoting the linear map of field extension $\F_{p^n}$
over $\F_p$.
\end{Definition}
Additionally we can define a subspace based on linear map $T$:
\begin{Definition}
A subspace $W\subset V$ is called {\bf $T$-invariant} when 
$$Tu \in W, ~~ \forall \text{ vector }u\in W$$

@@

We define the nullspace of a polynomial as:
\begin{Definition}[Nullspace of a polynomial]
\label{def:nullspace}
For any polynomial $g(x) \in \F[x]$, the null space of $g(T)$ consists of all vectors $u$ such that
$g(T)u = 0$.
\end{Definition}
Finally we get to the most important concept: $T$-Order, which derives the construction of a ring extended 
by a field requiring a minimal polynomial.
\begin{Definition}
For any vector $u \in V$, the monic polynomial $g(x) \in \F[x]$ with smallest degree such that
$g(T)u = 0$ is called the {\bf $T$-Order} of $u$ or {\bf minimal polynomial} of $u$.

@@

Using the concepts introduced above, along with other basic concepts in linear algebra and finite field theory,
we can derive the following theorems.

@@

Let $k$ be a polynomial whose degree is larger than that of $f(x)$, i.e., $deg(k(x))>deg(f(x))$, then
$$f(x)|k(x)~~\text{iff}~k(T)=0$$

@@

This equation denotes that $\beta$ is one of $p^n$ roots of polynomial $\Func(x) = \sum_if_i(x^{p^i})$.
However, the maximum number of roots allowed equals the degree of $\Func(x)$, which is 
$p^{n-1}<p^n$, we find a contradiction. Thus we conclude that
$$\textit{Both characteristic and minimal polynomial is }x^n-1$$
\end{Proof}

From the lemmas above we can deduce the following corollary:
\begin{Corollary}
\label{corol:1}
An element $\alpha \in \F_{p^n}$ is a {\bf normal element} if and only if $Ord_{\alpha,\sigma}(x) = x^n-1$.
\end{Corollary}
\begin{Proof}
Normal bases require $\{\beta,\beta^2,\dots,\beta^{2^{n-1}}\}$ to be linearly independent,
which is equivalent to the following expression:
$$\forall f(x)\in \F_p[x],~deg(f(x))<n$$
Furthermore, because there are no annihilators in the $T$-Order, this implies that
$$Ord_{\alpha,\sigma}(x) = x^n-1$$
\end{Proof}

@@

Ultimately, since there always exist at least one element fulfilling requirements in Theorem \ref{thm:1}
or Theorem \ref{thm:2}, we obtain the following theorem:
\begin{Theorem}[Normal Basis Theorem over Finite Fields]
There always exists a normal basis of $\F_{p^n}$ over $\F_p$.
\end{Theorem}
\section{Construction of General Normal Bases}
After proving the existence of normal bases, the upcoming question is how to find such a normal basis/element.
For general normal basis identification, there are two methods widely used: L\"uneburg's algorithm
and Lenstra's algorithm.

\subsection{L\"uneburg's Algorithm}
L\"uneburg's algorithm can be described with the following steps:

@@

Use Lemma \ref{lem:3}, the minimal/characteristic polynomial for $\sigma$ is $x^n-1$ implies that
any annihilator of $\alpha^i$ (i.e., $f_i(\sigma)$) divides $x^n-1$. Meanwhile, $\{\alpha^i~|~0<i<n-1\}$ forms a (standard) basis
of field $\F_{p^n}$, this means $\alpha_i$ are linearly independent with each other ($i=0,1,\dots,n-1$).
Since linear map $f(\sigma)$ guarantees that all elements in the field 
have order $lcm(f_0,\dots,f_{n-1})$, i.e., $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
By contradiction we can prove that there is no factor of $x^n-1$ that can be divided by the product 
of elements in set $\{f_0,f_1,\dots,f_{n-1}\}$. This actually corresponds to the assertion in the first step of the algorithm:
$$x^n-1 = lcm(f_0,f_1,\dots,f_{n-1})$$

@@

\subsection{Lenstra's Algorithm}
L\"uneburg's algorithm is an analytic method, which requires relatively high computational complexity 
mainly because of the factorization in its second step. To overcome the high cost, an inductive method 
is proposed, which is Lenstra's algorithm. It allows for setup of heuristics to accelerate the procedure.
Before introducing the details of the algorithm, we demonstrate two preliminary lemmas.
\begin{Lemma}
\label{lem:4}
For an arbitrary element $\theta \in \F_{p^n}$ that $Ord_\theta(x) \neq x^n - 1$,
let $g(x) = (x^n - 1)/Ord_\theta(x)$. There exists another element $\beta$ such that $g(\sigma)\beta = \theta$.
\end{Lemma}
\begin{Proof}
Assume $\gamma$ is the desired normal element. From the definition of normal element we derive 
$$\exists f(x)\in \F_{p^n}[x],~~f(\sigma)\gamma = \theta$$

@@

\begin{Lemma}
\label{lem:5}
Define $\theta$ and $g(x)$ as in Lemma \ref{lem:4}. Assume there exists a solution 
$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Respectively there exists a nonzero element $\eta$ such that
$$g(\sigma)\eta = 0$$
and
$$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$$
\end{Lemma}
\begin{Proof}
Assume $\gamma$ is the desired normal element. Similarly,
$$\exists \eta = Ord_\theta(\sigma)\gamma \neq 0,~~g(\sigma)\eta = 0$$

@@

Based on Lemmas \ref{lem:4} and \ref{lem:5}, Lenstra's algorithm is described below:

\begin{enumerate}[{1)}]
\item Take an arbitrary element $\theta \in \F_{p^n}$, determine $Ord_\theta(x)$.
\item If $Ord_\theta(x) = x^n - 1$ then algorithm terminates and return $\theta$ as a normal element.
\item Otherwise, compute $g(x) = (x^n - 1)/Ord_\theta(x)$, and solve $\beta$ from $g(\sigma)\beta = \theta$.
\item Determine $Ord_\beta(x)$. If $deg(Ord_\beta(x)) > deg(Ord_\theta(x))$ then replace $\theta$ by $\beta$ and go to 2);
otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a nonzero element $\eta$ such that $g(\sigma)\eta = 0$,
replace $\theta$ by $\theta + \eta$ and determine the order of new $\theta$, and go to 2).
\end{enumerate}

@@

Directly constructing $\lambda$-Matrix is difficult for general NBs. 
In Section \ref{sec:nbdesign}, we introduced the concept of a multiplication table (M-table) and explained how its 
entries are transformed from the $0$-th $\lambda$-Matrix. 
Recall the M-table is defined as
\begin{equation}
\label{eqn:appmt}
\beta
\begin{bmatrix}
\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
\end{bmatrix}
= {\bf T}
\begin{bmatrix}
\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
\end{bmatrix}
\end{equation}
Therefore we can compute M-table and then covert it to $\lambda$-Matrix.
Using Equation \ref{eqn:appmt}, we derive entry $T_{i,j}$ in the M-table by equating it to
the $j$-th bit of the NB representation of element $\beta\cdot\beta^{2^{i}}$.
For example, the $i$-th row of the M-table for NB in Example \ref{ex:Rijndael} satisfies
$$\beta\cdot\beta^{2^i} = \sum_{j=0}^{7} T_{i,j}\beta^{2^j}$$
Thus, the M-table can be written as
\begin{equation*}
T = \left(\begin{array}{lccccccr}
0 &1 &0 &0 &0 &0 &0 &1 \\
0 &1 &0 &0 &1 &1 &0 &1 \\
0 &0 &1 &0 &0 &1 &0 &1 \\
0 &0 &0 &0 &0 &0 &0 &1 \\
1 &1 &1 &1 &1 &1 &1 &1 \\
0 &0 &0 &0 &1 &0 &0 &0 \\
1 &1 &0 &1 &0 &1 &0 &0 \\
1 &0 &0 &1 &1 &0 &1 &0
\end{array}\right)
\end{equation*}
Using equation $M_{i,j}^{(0)} = T_{j-i,-i}$, we can obtain the
$\lambda$-Matrix in Equation \ref{eqn:mat_R}.

### 2017-03-03 dissertation/submission/appendix2.tex
after submission: revision 2

\chapter{Optimal Normal Basis}
\label{append:ONB}
The number of non-zero entries in $\lambda$-Matrix or multiplication table (M-table) is known as \emph{Complexity $(C_N)$}.
To define optimal normal basis, it is necessary to find the lower bound of $C_N$.
\begin{Theorem}
If $\mathcal{N}$ is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then non-zero entries in matrix $C_N\geq 2n-1$.
\end{Theorem}
\begin{Proof}
Let basis $\mathcal{N} = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, let us denote $\beta^{p^i}$ by $\beta_i$ for the sake of simplification.
Then $\sum_{i=0}^{n-1} \beta_i = trace\ \beta$. 
Denote $trace\ \beta$ by $b$, consider a $n\times n$ matrix $M^{(0)}$. Then
$$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$$

Therefore, the sum of all rows in $M^{(0)}$ is an n-tuple with $b$ as the first element and zeros elsewhere.
So the first column always includes at least one nonzero element. 
Since the sum of entries in each of other columns should equal to zero (modulo 2), they 
need to include an even number of ``1"s.
Meanwhile, in order to maintain the linear independence among each row, they cannot be all ``0"s. 
Therefore, there are always at least two non-zero elements in each column.

@@

ONBs are widely used in finite field circuit design, not only because its low complexity, 
but also because of the convenience to construct the $\lambda$-Matrix and M-table for them. 
To explain the convenience and the reason behind it, we introduce two types of ONBs and the method 
to construct them.
The following section refers to \cite{rosingbook}.
\section{Construction of Optimal Normal Basis}
{\bf Type-I ONB} over $F_{2^n}$ satisfies following criteria:
\begin{itemize}
\item $n+1$ must be prime.
\item 2 must be primitive in $\mathbb{Z}_{n+1}$.
\end{itemize}
The second criterion indicates that powers of 2 (exponent from 0 to $n-1$) modulo $n+1$ must cover all integers
from 1 to $n$.

In following, we derive a simple way to construct the $\lambda$-Matrix of type-I ONB 
from the criteria above. 
Assume $\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ from k-th $\lambda$-Matrix. Then the crossproduct
term can be written as
\begin{equation}
\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1} \lambda_{ij}^{(k)} \beta^{2^k}
\end{equation}
Suppose we only care about k=0. So simplified to following equations:

@@

Solution $(i,j)$ implies location of entries that equal to ``1" in $\lambda$-Matrix. 
Let $\beta$ be an optimal normal element, 
$\{\beta^{2^i}~|~0\leq i<n\}$ cover all powers of $\beta$ and generates the basis. Thus by solving
$$\begin{cases}
2^i + 2^j = 1 \pmod{n+1}\\
2^i + 2^j = 0 \pmod{n+1}
\end{cases}$$
we obtain $\lambda$-Matrix $M^{(0)}$. This construction method does not require information about 
primitive polynomial nor normal element, so it is very convenient for circuit designers.
Basis $\mathcal N_1$ in Example \ref{ex:app4} is a type-I ONB over $\F_{2^4}$.

@@

\section{Optimal Normal Basis Multiplier Design}
% Customized Galois Field IEEE1363-2000
Designers can easily generate the $\lambda$-Matrix for GF multipliers, which is sufficient to derive the 
structure of the circuit. However, our GB based approach requires specifying the exact normal element $\beta$,
{\it i.e. } it is necessary to obtain $t$ as $\beta = \alpha^t$ before executing our technique.
L\"uneburg's algorithm and Lenstra's algorithm do not guarantee the output normal element is optimal 
normal element, and usually result in high computation cost. Actually if the type of ONB within the design 
is known, instead of looking up the optimal normal element, we can construct a special irreducible $p(\alpha)$
such that optimal normal element equals to the primitive element: $\beta=\alpha$. The concepts and algorithms 
refer to IEEE standard 1363-2000 \cite{IEEE1363}.

@@

For type-II ONB, the following iterative algorithm is required to generate desired 
irreducible polynomial:

>>>>

\chapter{Optimal Normal Basis}
\label{append:ONB}
The number of nonzero entries in $\lambda$-Matrix or multiplication table (M-table) is known as \emph{Complexity $(C_N)$}.
To define optimal normal basis, it is necessary to find the lower bound of $C_N$.
\begin{Theorem}
If $\mathcal{N}$ is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then nonzero entries in matrix $C_N\geq 2n-1$.
\end{Theorem}
\begin{Proof}
Let basis $\mathcal{N} = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, let us denote $\beta^{p^i}$ by $\beta_i$ for the sake of simplification.
Then $\sum_{i=0}^{n-1} \beta_i = trace\ \beta$. 
Denote $trace\ \beta$ by $b$, consider a $n\times n$ matrix $M^{(0)}$. Then
$$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$$

Therefore, the sum of all rows in $M^{(0)}$ is an n-tuple with $b$ as the first element and zeros elsewhere.
So the first column always includes at least one nonzero element. 
Since the sum of entries in each of the other columns should equal zero (modulo 2), they 
need to include an even number of ``1"s.
Meanwhile, in order to maintain the linear independence among each row, they cannot be all ``0"s. 
Therefore, there are always at least two nonzero elements in each column.

@@

ONBs are widely used in finite field circuit design, not only because of its low complexity, 
but also because of the convenience to construct the $\lambda$-Matrix and M-table for them. 
To explain the convenience and the reason behind it, we introduce two types of ONBs and the method 
to construct them.
The following section refers to \cite{rosingbook}.
\section{Construction of Optimal Normal Basis}
{\bf Type-I ONB} over $F_{2^n}$ satisfies following criteria:
\begin{itemize}
\item $n+1$ must be prime.
\item 2 must be primitive in $\mathbb{Z}_{n+1}$.
\end{itemize}
The second criterion indicates that powers of 2 (exponent from 0 to $n-1$) modulo $n+1$ must cover all integers
from 1 to $n$.

In the following, we derive a simple way to construct the $\lambda$-Matrix of type-I ONB 
from the criteria above. 
Assume $\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ from k-th $\lambda$-Matrix. Then the crossproduct
term can be written as
\begin{equation}
\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1} \lambda_{ij}^{(k)} \beta^{2^k}
\end{equation}
Suppose we only care about $k=0$. So simplified to following equations:

@@

Solution $(i,j)$ implies location of entries that equal ``1" in $\lambda$-Matrix. 
Let $\beta$ be an optimal normal element, 
$\{\beta^{2^i}~|~0\leq i<n\}$ cover all powers of $\beta$ and generates the basis. Thus by solving
$$\begin{cases}
2^i + 2^j = 1 \pmod{n+1}\\
2^i + 2^j = 0 \pmod{n+1}
\end{cases}$$
we obtain $\lambda$-Matrix $M^{(0)}$. This construction method does not require information about 
primitive polynomial or normal element, so it is very convenient for circuit designers.
Basis $\mathcal N_1$ in Example \ref{ex:app4} is a type-I ONB over $\F_{2^4}$.

@@

\section{Optimal Normal Basis Multiplier Design}
% Customized Galois Field IEEE1363-2000
Designers can easily generate the $\lambda$-Matrix for GF multipliers, which is sufficient to derive the 
structure of the circuit. However, our GB based approach requires specifying the exact normal element $\beta$,
i.e., it is necessary to obtain $t$ as $\beta = \alpha^t$ before executing our technique.
L\"uneburg's algorithm and Lenstra's algorithm do not guarantee the output normal element is the optimal 
normal element, and usually result in high computation cost. Actually if the type of ONB within the design 
is known, instead of looking up the optimal normal element, we can construct a special irreducible $p(\alpha)$
such that the optimal normal element is equal to the primitive element: $\beta=\alpha$. The concepts and algorithms 
refer to IEEE standard 1363-2000 \cite{IEEE1363}.

@@

For type-II ONB, the following iterative algorithm is required to generate the desired 
irreducible polynomial:
